### 2017-06-06 sheets/talk-6/sheet-6.tex
Sheet 6: Rewording, Removal of Exercise 3, 7 and 8

\hint{Read the slides for the definitions of the six constructions}

@@

\begin{exercise}[Product is unique up-to isomorphism]
  Proof that for every category $\cat{A}$ all products are isomorphic. Thus we can speak of \emph{the} product.
\end{exercise}

@@

\begin{exercise}[Greatest lower bound]
  Define the \emph{greatest lower bound} for an family of objects in a poset and proof that it is the arbitrary product of the family when the poset is seen as a category.
\end{exercise}

@@

\begin{exercise}[Least upper bound]
  Define the \emph{least upper bound} for a class of objects in a poset and proof that it is the arbitrary binary coproduct of the poset seen as a category.
\end{exercise}

\begin{answer}
  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
  A \demph{upper bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: x_i \le z$.
  $z \in A$ is a \demph{least upper bound} of this family if $\forall a \in A: \text{$a$ is a upper bound} \Rightarrow z \le a$.

  Let $z$ be a least upper bound of the family $(x_i)_{i\in I}$. We show that $z$ is a coproduct of $(x_i)_{i\in I}$.
  \begin{proof}
    We first have to define a family of maps $\Bigl(x_i \toby{p_i} z\Bigr)_{i \in I}$.
    This is trivial since we have $\forall i \in I: x_i \le a$.
    Now let $a \in A$ be and objects and let $\Bigl(x_i \toby{f_i} a\Bigr)_{i \in I}$ be a family of maps to $a$.
    The existing of this maps means that $a$ is a lower bound of $(x_i)_{i\in I}$.
    Now by definition of least lower bound, $z \le a$, thus we have a arrow $\bar{f} \from z \to a$.
    All arrows are unique, thus $\bar{f}$ is unique and the diagram commutes. \qedhere
  \end{proof}
\end{answer}

@@

\begin{definition}{Minus-Core of Linear Maps in Vector Spaces}
  Let $V, W$ be Vector Spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
  We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
  Remember the definition $\forall g \from V \to W: core(g) := \set{ v \in V | g(v) = 0 }$.
  Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
\end{definition}

@@

\begin{exercise}[Coequalizer of sets]
  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
  \begin{itemize}
    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
      \begin{itemize}
        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
        \item[(iv)]Now show that $S$ is transitive.
          \begin{hint}
            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
          \end{hint}
        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
      \end{itemize}
    \item[(b)]Now we will define the coequalizer of $\Set$.
      Let $X, Y \in \Set, s, t \from X \to Y$.
      We make the following definitions:
      \begin{alignat*}{2}
        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
        C    &:=& Y / \sim
      \end{alignat*}
      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
      Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
      We define $i \from Y \to E$ with $i(y) := [y]$.
      \begin{itemize}
        \item[(i)]Show that the following diagramm commutes:
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
        \item[(ii)]Let
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
          }$
          be a cummutative diagramm. Show that $f$ respects equivalence, i.\,e. $\forall (y_1, y_2) \in \sim: f(y_1) = f(y_2)$.
          \begin{hint}
            ``Destruct'' $(y_1, y_2) \in \sim$ and do a $\nat$-induction.
          \end{hint}
        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
          \[ \xymatrix{
            & & A  \\
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
            & & C \ar@{.>}[uu]|{\bar{f}}
          } \]
        \item[(iv)]Show that the fork
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
          is the coequalizer of
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
          }$.
      \end{itemize}
  \end{itemize}
\end{exercise}

@@

\begin{exercise}[Pullbacks\,vs.\,producs]
  Show that if $Z$ is terminal and
  \[ \xymatrix{
    P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
    X \ar[r]^s & Z
  } \]
  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$ then $X, Y$ is the product in $\cat{A}$.
\end{exercise}

\begin{exercise}[Pushouts\,vs.\,coproducs]
  Show that if $Z$ is initial and
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \\
    Y \ar[r]^{p_2} & P
  } \]
  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$, then $X + Y$ is the coproduct in $\cat{A}$.
\end{exercise}

>>>>

\hint{Read the slides for the definitions of the six constructions.}

@@

\begin{exercise}[Product is unique up-to isomorphism]
  Prove that for every category $\cat{A}$ and every objects $X, Y \in \cat{A}$
  all products of $X$ and $Y$ are isomorphic. Thus we can speak of \emph{the} product of $X$ and $Y$.
\end{exercise}

@@

\begin{exercise}[Greatest lower bound]
  Define the \emph{greatest lower bound} for a family of objects in a poset and prove that it is the arbitrary product of the family when the poset is seen as a category.
\end{exercise}

@@

@@

\begin{definition}{Minus-Core of Linear Maps in Vector Spaces}
  Let $V, W$ be Vector Spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
  We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
  Remember that for $g \from V \to W$ we have defined $core(g) := \set{ v \in V | g(v) = 0 }$.
  Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
\end{definition}

@@

@@

\begin{exercise}[Pushouts\,vs.\,coproducts]
  Show that if $Z$ is initial and the pullback square
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \\
    Y \ar[r]^{p_2} & P
  } \]
  is the pullback of
  $ \xymatrix{
    X & Z \ar[l]_s \ar[r]^t & Y \\
  } $
  then $P$ is the coproduct of $X$ and $Y$.

  Remark: Note that the dual statement holds for pullbacks of terminal objects and products.
\end{exercise}
